---
title: "Assignment1"
author: "Jaymo Kim"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section 4.1.1
#### Exercise 2
#### Part(a)
```{r}
Fib <- numeric(30)
Fib[1] <- Fib[2] <- 1
for(i in 3:30) {
    Fib[i] <- Fib[i - 2] + Fib[i - 1]
}
Fib.ratio <- numeric(29)
for(i in 2:30) {
    Fib.ratio[i] <- Fib[i]/Fib[i - 1]
}
Fib.ratio
```

The sequence of ratios of Fibonacci numbers seems to be converging to `r Fib.ratio[29]`.

#### Part(b)
```{r}
gr <- (1 + sqrt(5))/2
```

The golden ratio is equal to `r gr` and the sequence computed above converges to this golden ratio.
***
Proof:
***
Let $\phi$ denote the golden ratio $= \frac{1 + \sqrt{5}}{2}$.
Then $\phi$ has a property of $\phi = 1 + \frac{1}{\phi}$.
Now, let $R_n = \frac{F_{n+1}}{F_n}$, where $F_n$ denotes the $n^{th}$ Fibonacci number for n = 1, 2,...
By definition, we can derive that $R_n = \frac{F_{n+1}}{F_n} = \frac{F_n + F_{n-1}}{F_n} = 1 + \frac{1}{R_{n-1}}$.
Then, $\forall n = 1, 2, 3,\dots$, $|R_n - \phi| = |\left(1 + \frac{1}{R_{n-1}}\right) - \left(1 + \frac{1}{\phi}\right)| = |\frac{1}{R_{n-1}} - \frac{1}{\phi}| = |\frac{\phi - R_{n-1}}{R_{n-1}\phi}| \leq \frac{1}{\phi}|R_{n-1} - \phi| \leq \left(\frac{1}{\phi}\right)^{n-1}|R_1 - \phi|$.
Since $0 < \frac{1}{\phi} < 1$, $\lim_\limits{n \to \infty} \left(\frac{1}{\phi}\right)^{n-1} = 0$ which implies $\lim_\limits{n \to \infty} |R_n - \phi| = 0$.
Therefore, $\lim_\limits{n \to \infty} \frac{F_{n+1}}{F_n} = \phi$.

#### Exercise 3
#### Part(a)
Answer is summing up all values from 1 to 5, $1+2+3+4+5 = 15$
```{r}
answer <- 0 
for(j in 1:5) answer <- answer + j
answer
```

#### Part(b)
Answer is a list consisting of all values of j (1 through 5) in an increasing order.
```{r}
answer <- NULL
for(j in 1:5) answer <- c(answer,j)
answer
```

#### Part(c)
Similar to part b, answer is a listing consisting of all values of j in an increasing order with 0 in the beginning of the list.
```{r}
answer <- 0
for(j in 1:5) answer <- c(answer,j)
answer
```

#### Part(d)
Answer is obtained by multiplying 1 through 5, $1*2*3*4*5 = 120$.
```{r}
answer <- 1
for(j in 1:5) answer <- answer * j
answer
```

#### Part(e)
Answer would be a list containing 3 as a first element and following elements to be 7 * (previous element in the list) %% 31.
Without knowing the rule used to determine this sequence, it would have been extremely hard to predict successive elements as the sequence requires a knowledge of calling the element in a list.
```{r}
answer <- 3
for(j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
answer
```

### Section 4.1.2
#### Exercise 4
```{r}
interest <- function(p, n) {
    if(n > 3) {
        i = 0.05
    }
    else {
        i = 0.04
    }
    i = p*((1 + i)^n - 1)
    return(i)
}
```

#### Exercise 5
```{r}
mortgage <- function(n, P, open) {
    if(open == TRUE) {
        i = 0.005
    }
    else {
        i = 0.004
    }
    R = (P*i)/(1-(1+i)^(-n))
    return(R)
}
```

### Section 4.1.3
#### Exercise 2
```{r}
Fibonacci <- c(1, 1)
# I will condition the Fibonacci numbers to show until the lateset Fibonacci number is less than 300 (preventing infinite loop).
while(Fibonacci[length(Fibonacci)] < 300){
    Fibonacci <- c(Fibonacci, Fibonacci[length(Fibonacci)])
    Fibonacci[length(Fibonacci)] <- Fibonacci[length(Fibonacci) - 1] + Fibonacci[length(Fibonacci) - 2]
}
Fibonacci <- Fibonacci[-length(Fibonacci)] # This is to get rid of the latest Fibonacci number which is greater than 300, or upper bound.
```

#### Exercise 4
```{r}
i <- 0.006
j <- 1
while(j > 0.000001) {
    newi <- (1 - (1 + i)^(-20))/19
    j <- abs(newi - i)
    i <- newi
}
i
```
Even if you start with different starting guesses, the value of interest rate, i, converges to `r i`.

#### Exercise 5
```{r}
i <- 0.006
j <- 1
n <- 0
while(j > 0.000001) {
    newi <- (1 - (1 + i)^(-20))/19
    j <- abs(newi - i)
    i <- newi
    n <- n + 1
}
i
n
```
It can be noticed from this while loop that the number of iteration for a while loop differs based on the value of starting guesses.

### Section 4.1.5
#### Exercise 2
#### Part(a)
```{r}
Eratothenes <- function(n) {
    # Print prime numbers up to n (based on the sieve of Eratothenes)
    if(n >= 2) {
        sieve <- seq(2, n)
        primes <- c()
        while(length(sieve) > 0) {
            p <- sieve[1]
            primes <- c(primes, p)  # Saving the first element of sieve in a list of prime numbers.
            sieve <- sieve[(sieve %% p) != 0]   # Eliminating all the multiples of p, or the first element in sieve.
        }
        return(primes)
    } else {
        stop("Input value of n should be at least 2.")  # The first prime number is 2.
    }
}
```

The while loop takes the first element in the list sieve and designate the element as a prime number.
Then the loop eliminates all the multiples of the first element in the sieve so that the first element after elimination can be a prime number again.

#### Part(b)
Once $p >= sqrt(n)$ all remaining entries in $sieve$ are prime because all the other entries in $sieve$ that are higher than $p$ but are non-prime numbers have been already eliminated for they are multiples of prime numbers smaller or equal to p.
For example, let $n = 10$ and $p \geq \sqrt{n} = 4$. Then, $sieve = (2, 3, \not{4}, 5, \not{6}, 7, \not{8}, \not{9}, \not{10})$. All the non-prime numbers greater than $p$, (6, 8, 9, 10), have already been eliminated for they are multiples of 2, 3, 4, and 5. And the remaining elements in $sieve$ that are greater than $p$, 5 and 7, are the prime numbers.

#### Part(c)
```{r}
Eratosthenes <- function(n) {
    if (n >= 2) {
        sieve <- seq(2, n)
        primes <- c()
        repeat {
            p <- sieve[1]
            primes <- c(primes, p)
            sieve <- sieve[(sieve %% p) != 0]
            if (p >= sqrt(n)) {
                primes <- c(primes, sieve)
                break
            }
        }
        return(primes)
    } else {
        stop("Input value of n should be at least 2.")
    }
}
```

### Section 4.2.1
#### Exercise 2
#### Part(a)
```{r}
compound.interest <- function(P, i.r, n) {
    interest <- P*(1 + i.r)^n
    return(interest)
}
```

#### Part(b)
In this situation, $P = \${1000}, i.r = 1\% per month, n = 30 months$.
Then, at the end of the 30 month period, Mr. Ng will have $\$$`r compound.interest(P=1000, i.r=0.01, n=30)`.

#### Exercise 3
```{r}
# The function bisection have five arguments (a, b, func, nmax, tol) where a and b denote the two endpoints, func denote a user-supplied function, nmax denotes the number of maximum iteration in order to prevent an infinite loop, and tol for a level of tolerance .
bisection <- function(a, b, func, nmax, tol) {
    n <- 1
    while(n <= nmax) {
        c <- (a + b)/2
        if(func(c) == 0 || (b - a)/2 < tol) return(c)
        n <- n + 1
        if(sign(c) == sign(a)) a <- c
        else b <- c
    }
    return("Method failed.")
}
```

### Section 4.4.1
#### Exercise 1
```{r}
mergesort <- function(x, decreasing) {
    len <- length(x)
    if(len < 2) result <- x
    else {
        if(decreasing == TRUE) {
            y <- x[1 : (len / 2)]
            z <- x[(len / 2 + 1):len]
            result <- c()
            y <- mergesort(y, decreasing)
            z <- mergesort(z, decreasing)
            while(min(length(y), length(z)) > 0) {
                if(y[1] > z[1]) {
                    result <- c(result, y[1])
                    y <- y[-1]
                } else {
                    result <- c(result, z[1])
                    z <- z[-1]
                }
            }
            if(length(y) > 0) result <- c(result, y)
            else result <- c(result, z)
        }
        else {
            y <- x[1 : (len / 2)]
            z <- x[(len / 2 + 1):len]
            result <- c()
            y <- mergesort(y, decreasing)
            z <- mergesort(z, decreasing)
            while(min(length(y), length(z)) > 0) {
                if(y[1] < z[1]) {
                    result <- c(result, y[1])
                    y <- y[-1]
                } else {
                    result <- c(result, z[1])
                    z <- z[-1]
                }
            }
            if(length(y) > 0) result <- c(result, y)
            else result <- c(result, z)
        }
    }
    return(result)
}
```

#### Exercise 2
#### Part(a)
```{r}
# install.packages("Deriv")
library(Deriv)
newton <- function(f, g, x0, y0, TOL) {
    fx <- function(f, x, y) {
        foo <- eval(Deriv(f, "x"))
        return(foo(x, y))
    }
    fy <- function(f, x, y) {
        foo <- eval(Deriv(f, "y"))
        return(foo(x, y))
    }
    gx <- function(g, x, y) {
        foo <- eval(Deriv(g, "x"))
        return(foo(x, y))
    }
    gy <- function(g, x, y) {
        foo <- eval(Deriv(g, "y"))
        return(foo(x, y))
    }
    f1 <- f(x0, y0)
    g1 <- g(x0, y0)
    while(abs(f1) > TOL || abs(g1) > TOL) {
        dn <- fx(f, x0, y0) * gy(g, x0, y0) - fy(f, x0, y0) * gx(g, x0, y0)
        x1 <- x0 - (gy(g, x0, y0) * f(x0, y0) - fy(f, x0, y0) * g(x0, y0))/dn
        y1 <- y0 - (fx(f, x0, y0) * g(x0, y0) - gx(g, x0, y0) * f(x0, y0))/dn
        f1 <- f(x1, y1)
        g1 <- g(x1, y1)
        x0 <- x1
        y0 <- y1
        
    }
    return(c(x0, y0, f1, g1))
}
```

#### Part(b)
```{r}
f <- function(x, y) x + y
g <- function(x, y) x^2 + 2*y^2 - 2
newton(f, g, x0 = -2, y0 = 4, TOL = 0.000001)
```


### Chapter 4
#### Exercise 1
```{r}
directpoly <- function(x, c) {
    p <- 0
    n <- length(c)
    for (i in 1 : n) {
        p <- p + c[i]*x^(i-1)
    }
    return(p)
}
```

#### Exercise 2
```{r}
hornerpoly <- function(x, c) {
    n <- length(c)
    a <- matrix(0, nrow = length(x), ncol = length(c))
    a[, n] <- c[n]
    for(i in (n - 1):1) {
        a[, i] <- a[, i+1] * x + c[i]
    }
    return(a[, 1])
}
```

#### Exercise 3
#### Part(a)
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000), 
                       c(1, -2, 2, 3, 4, 6, 7)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), 
                       c(1, -2, 2, 3, 4, 6, 7)))
```
Time elapsed for the directpoly function is about 2.50 while time elapsed for the hornerpoly function is about 0.97.
Therefore, we can conclude that using the hornerpoly function is much faster compared to the directpoly function.

#### Part(b)
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000), 
                       c(-3, 17, 2)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), 
                       c(-3, 17, 2)))
```
Time elapse for the directpoly function is about 0.58 while time elapsed for the hornerpoly function is about 0.32.
Although using the hornerpoly function is much faster, the time difference has been reduced by alot when the number of polynomial coefficients is smaller.